
```{r setup, echo = F}
library(knitr)
opts_knit$set(out.format = 'html')
opts_chunk$set(highlight = TRUE, tidy = FALSE)
render_html()
```

# R 

Programming Literate and Reproducible Research

John Painter

# Visualizating Data

- The larger the data, the greater the need to see it in pictures
- Challenge to visualize data with 'traditional' software
 * SAS has limited options
 * Excel poor choice for data management
 * Propriatary software expensive (ArcGis) and limited (ArcGis=maps only)
- Some of best visualizations done by newer (and cheaper) software

# New York Times

- The staff of New York Times info graphics department is...[video please](http://vimeo.com/21020824)

- NYT [blog](http://chartsnthings.tumblr.com/) about how they brainstorm and design effective graphics

- Former intern has popular book and web-site on visualizations, [Flowing Data](http://flowingdata.com/) 

- Basic workflow is [R](http://www.r-project.org/) for data and chart, Adobe for post chart clean up, and convert to [D3 (d3js.org)](http://d3js.org/) for interactive, web-based version

# R is
- Free
- More advanced than SAS, STATA, et al.
- Better visualizations
- Better for large data sets (CERN)

# Free

- IBM wrote first code, "S", and published it for anyone to use
- S-plus licensed a version
- Two guys in NZ (Ross I. and Robert G.) published a free version, "R"
- Add-on code downloaded as needed from 87 reposisitories (17 in US)
- [RStudio (rstudio.org)](http://rstudio.org/) for [R (r-project.org)](http://www.r-project.org/) provides GUI for organizing projects, files, and plots

# Advanced

- Around 3,500 specialized packages
- Authors are stats profressors, engineers, end-users
- Becoming a primary software for academic biostatistics, like [Vanderbilt](http://biostat.mc.vanderbilt.edu/wiki/Main/RS)

# Visualizations
- Wilkinson and Tufte
- 'Grammar of Graphics' 
- Examples online are motivational
 * [Mariano Rivera's saves](http://www.nytimes.com/interactive/2012/05/05/sports/baseball/mariano-rivera-and-his-peers.html?ref=baseball)
 * [Maritime trade routes](http://spatialanalysis.co.uk/2012/03/mapped-british-shipping-1750-1800/)
 * [London cycle journey](http://spatialanalysis.co.uk/2012/02/great-maps-ggplot2/)
 * [Wind patterns](http://hint.fm/wind/)
 
# GoogleMaps from R data 
## The map below is an example of how to send data through google and get a dynamic map back.  Note:  need internet connection to get map (a disadvantage compared with D3 interactive images)
```{r ,results='asis', echo=FALSE, message=FALSE}
require(googleVis)
require(XML)
states <- data.frame(state.name, state.x77)
G3 <- gvisGeoMap(states, "state.name", "Illiteracy",
                 options=list(region="US", dataMode="regions",
         width=600, height=400))
print(G3, "chart")
``` 

# Reproducible Research

- Prime focus of international R conference, [UserR! 2012](http://biostat.mc.vanderbilt.edu/wiki/Main/UseR-2012)
- [Criteria for Reproducible Research](http://www.reproducibleresearch.net/index.php/Main_Page)
- Literate programming: 
  * Documentation (and perhaps a statistical report) and code are maintained in one file *
- An example?  *This* presentation is written in R, using [knitr](http://yihui.name/en/2012/05/how-to-make-html5-slides-with-knitr/) to make HTML5 slides with [Slidy](http://www.w3.org/Talks/Tools/Slidy2/#(1))

# HTML5 slides
- Several programs generate slides from a generic markup file (.md)
- R creates the .md file
- PanDoc converts to 

 -- dzslides 
  ---pandoc -s -S -i -t dzslides --mathjax DNA-Presentation.md -o DNA-Presentation.html

 -- Slidy
   ---pandoc -s -S -i -t slidy --mathjax R.md -o R-slidy.html
   --- OR
    ---- library(slidify)
    ---- slidify("DNA-Presentation.Rmd")


